---
title: "R analysis file"
author: "Mousam"
date: "November 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

loading library, 'dplyr' for data manipulation and 'ggplot2' for data visualization
```{r}
library(dplyr)
library(ggplot2)
```
loading the data from the dataset file in csv format
```{r}
wage <- read.csv("Wage.csv", stringsAsFactors = FALSE)
#stringsAsFactors is set false to avoud crating factors if value is string
#write.csv(wage , "Wage.csv") to write data into a csv file
#wage
```
the unnecessary index column is removed
```{r}
wage<- select(wage , -X)
#wage
```
other files like spss , excel and json can be loaded which need 'foreign', 'readxl' and 'rjson' respectively
```{r}
dim(wage) #dimensions of data
cat("No. of rows:" , nrow(wage)) #concatenating the string with number of rows
cat("\nNo. of column:", ncol(wage))
```

```{r}
names(wage)
colnames(wage)#both of these show the name of header of columns
#row.names(wage) would give name of row if there were. 
```

```{r}
head(wage)#shows first data
```
```{r}
tail(wage, 3)#second parameter specifies the number of examples to be displayed
```
To show the whole dataset in a new tab
```{r}
View(wage)
```
To count the factor values:
```{r}
table(wage$education)#table function counts the factor values
```
To change the name of values, to make name of values more uniform
```{r}
wage$education[wage$education=="5. Advanced Degree"] <- "Advanced Degree"
```
is.na means not available values, so we can clean and remove missing and unwanted data.
```{r}
sum(is.na(wage))
```
Here 55 means there are 55 cells in which data is missing.
```{r}
table(is.na(wage))
```
Colsums() sums the value along the column
```{r}
colSums(is.na(wage))
```
Here, health_ins and education are qualitative data and we will impute NA with mode value
```{r}
wage$education[is.na(wage$education)]<- names(table(wage$education))[table(wage$education) == max(table(wage$education))]
wage$health_ins[is.na(wage$health_ins)] <- names(table(wage$health_ins))[table(wage$health_ins) == max(table(wage$health_ins))]
```
wage is quantitative data, so we will replace NA with median value of the wage
```{r}
wage$wage[is.na(wage$wage)]<- median(wage$wage , na.rm = TRUE)
```
since the whole data was taken from Mid- atlantic region , we can remove it from the data frame.
```{r}
wage<- select(wage, -region)
```
```{r}
colSums(is.na(wage))
```

To replace the factor value health_ins , which has 1 or 2 
```{r}
wage$health_ins[wage$health_ins == "2. No"] <- "NO"
wage$health_ins[wage$health_ins == "1. Yes"] <- "Yes"
```
<h2> Analysis</h2>
```{r}
max(wage$wage) #gives maximum value 
min(wage$age) #gives minimum value
length(wage$year) #gives length of the vector
mean(wage$wage) #gives mean
median(wage$age) #gives median 
names(table(wage$race))[table(wage$race) == max(table(wage$race))] # gives mode
range(wage$wage) #gives range, i.e- maximum and minimum value
quantile(wage$age , c(0.0 , 0.25 , 0.50 , 0.75, 1)) #gives quantiles or quartiles
var(wage$wage)# gives variance
sd(wage$wage)# gives standard deviation
```

Or we can use summary() function to show various statistical calculations on dataset
```{r}
summary(wage)
```
str() function displays the internal structure of an R object
```{r}
str(wage)
```

Relation between variables:
```{r}
#correlation to show how strongly pair of variables (quantitative) are related.
cor(wage%>% select(age,logwage,wage))
```

correlation function can not be used on qualitative variables. We use chi square test for this purpose
```{r}
#Hypothesis testing
chisq.test(table(wage$jobclass,wage$education))
```
The p-value is less than the cut off value i.e 0.05. We will reject the null hypothesis and accept alternate hypothesis. That means jobclass and education are related.

<h2>Exploratory analysis</h2>
```{r}
g<- ggplot(data = wage , aes(wage))
g+geom_density()
```
```{r}
#this code generates normal distribution
library(moments)
normal<-as.data.frame(rnorm(500))
names(normal)<- c("Norm")
ggplot(normal)+geom_density(aes(Norm))

```
```{r}
skewness(normal)
```
```{r}
g<- ggplot(data = wage , aes(age))
g+geom_histogram()
```

```{r}
skewness(wage$age)
kurtosis(wage$age)
```
The skewness value of age shows that age is Positively skewed distribution or right skewed and the kurtosis values tells that its Leptokurtic distribution.

<h3>Bar plots</h3>

```{r}
ggplot(as.data.frame(table(wage$race)),aes(Var1,Freq)) +geom_col(aes(fill=Var1))
```

<h4>Lineplot</h4>
```{r}
ggplot(as.data.frame(table(wage$age),stringsAsFactors = FALSE) ,aes(Var1,Freq)) +geom_point()+geom_line(aes(group=1))+scale_x_discrete(breaks = c(18,28,38,48,58,68,80))
```

<h3>Boxplot</h3>
To show the quantiles and outliers in a single plot.
```{r}
boxplot(wage%>% select(age,wage))
```

<h2>Visualizations</h2>
```{r}
maritl<- as.data.frame(table(wage$maritl,wage$age), stringsAsFactors = FALSE)%>% group_by(Var1)
```
```{r}
g<- ggplot(maritl,aes(Var2,Freq))+geom_line(aes(group=Var1,color=Var1))
g + theme(axis.text.x = element_text(angle = 90 , hjust = 1))
```

```{r}
g<- g+scale_x_discrete(breaks = c(18,28,38,48,58,68,80))+ggtitle("Marital status and the age of the worker") +xlab("Age") + ylab("Count")
g+scale_color_discrete(name="Marital Status")
```
 This lineplot shows the maritial status of the workers with colour and count of worker with differnt age .
```{r}
ggplot(as.data.frame(table(wage$year)), aes(Var1,Freq)) + geom_col() + ggtitle("Count of the people in survey over the year")+xlab("Years") +ylab("Count")

```
 This barplot shows the number of data collected over the year.Maximum data was collected in year 2003.
```{r}
race <- as_tibble(table(wage$race))
```
Tibble is modern and efficient version of  the old dataframe.
```{r}
head(race,3)
```
```{r}
names(race) <- c("Race" , "Count")
names(race)
```
```{r}
g<- ggplot(race , aes(Race, Count))
g+geom_col(aes(fill = Race)) +labs(title = "Workers count on the basis of race")+ theme(legend.position = "")+ geom_text(aes(Race , Count , label=Count))
```
This plots shows the data has more white race workers in worksforce.
```{r}
edu <- as_tibble(table(wage$education))
names(edu) <- c("Education" ,"Count")
g<- ggplot(edu , aes(Education, Count))
g+geom_col(aes(fill = Education)) +ggtitle("Count off Workers on basis of Education") + theme(legend.position = "") +geom_text(aes(Education,Count,label=Count))
```
 This barplot shows that maximum number of workers are with HS grad.
```{r}
table(wage$education)
```
```{r}
health_age<-as_tibble(table(wage$age,wage$health_ins))
names(health_age)<- c("Age","Health_insurance","Count")
names(health_age)
```

In the plot below we are ploting the count of workers on basis of Age and Health Insurance.
```{r}
g<- ggplot(health_age,aes(Age,Count))
g+geom_line(aes(group=Health_insurance,color=Health_insurance))+geom_point(aes(color=Health_insurance))+scale_x_discrete(breaks=c(18,28,38,48,58,68,80))
```
 We can see that less number of worker with age 18 to 24 has health insurance.The people from 38 to 48 has maximum number of health insurance .
```{r}
ggplot(wage,aes(age)) +geom_density()
```
 This density plot shows that the maximum number of workers are from 38 to 48 group .
```{r}
indus_edu_hea <- as_tibble(table(wage$jobclass,wage$education,wage$health))
names(indus_edu_hea) <- c("Job","Education" , "Health" , "Count")
```
```{r}
g<- ggplot(indus_edu_hea , aes(Education, Count))
g+geom_col(aes(fill = Education)) +facet_grid(.~Health)+theme(axis.text.x = element_blank()) +labs(title = "Health Conditions and Education of workers")
```
 This barplot shows multiple type of data in single plot.This barplot is divided into two group on basis of health of workers and the color define the colors.
```{r}
g<- ggplot(indus_edu_hea, aes(Education, Count))
g+geom_col(aes(fill=Education)) +facet_grid(.~Job) +theme(axis.text.x = element_blank())+labs(title = "Education and the Job worker.")
```
 This plot shows that the lots of worker with HS grad are working in industry. 








