---
title: "R analysis file"
author: "Mousam"
date: "November 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#loading library, 'dplyr' for data manipulation and 'ggplot2' for data visualization
```{r}
library(dplyr)
library(ggplot2)
```
#loading the data from the dataset file in csv format
```{r}
wage <- read.csv("Wage.csv", stringsAsFactors = FALSE)
#stringsAsFactors is set false to avoud crating factors if value is string
#write.csv(wage , "Wage.csv") to write data into a csv file
wage
```
# the unnecessary index column is removed
```{r}
wage<- select(wage , -X)
wage
```
#other files like spss , excel and json can be loaded which need 'foreign', 'readxl' and 'rjson' respectively
```{r}
dim(wage) #dimensions of data
cat("No. of rows:" , nrow(wage)) #concatenating the string with number of rows
cat("\nNo. of column:", ncol(wage))
```
#
```{r}
names(wage)
colnames(wage)#both of these show the name of header of columns
#row.names(wage) would give name of row if there were. 
```

```{r}
head(wage)#shows first data
```
```{r}
tail(wage, 3)#second parameter specifies the number of examples to be displayed
```
#To show the whole dataset in a new tab
```{r}
View(wage)
```
#To count the factor values:
```{r}
table(wage$education)#table function counts the factor values
```
#To change the name of values, to make name of values more uniform
```{r}
wage$education[wage$education=="5. Advanced Degree"] <- "Advanced Degree"
```
#is.na means not available values, so we can clean and remove missing and unwanted data.
```{r}
sum(is.na(wage))
```
# Here 55 means there are 55 cells in which data is missing.















